---
title: "AssessmentPart1"
output: html_notebook
---

```{r}
# Check working path and files
getwd()
list.files(getwd())
```

```{r}
install.packages(c("sf", "tidyverse", "tmap", "tmaptools", "geojsonio", "sp", "reshape2", "stplanr", "leaflet", "broom", "maptools", "stplanr"))
library(tidyverse)
library(sf)
library(tmap)
library(tmaptools)
library(geojsonio)
library(sp)
library(reshape2)
library(stplanr)
library(leaflet)
library(broom)
library(tidyverse)
library(datasets)
library(graphics)
library(grDevices)
library(methods)
library(RColorBrewer)
library(classInt)
library(rgeos)
library(rgdal)
library(geojsonio)
library(classInt)
```
```{r fig.width=1, fig.height=10, echo=FALSE}
library(png)
library(knitr)
fig1_path <- "Birmingham.png"
figure1 <- readPNG(fig1_path, native = TRUE, info = TRUE)
include_graphics(fig1_path)
```

```{r}
#read the shapefile into a simple features object
BirminghamMapMSOA <- read_shape("D:\\GIS\\datas\\Birmingham_CDRC_2011_CensusGeodatapack\\shapefiles\\Birmingham_msoa11.shp", as.sf = TRUE)
```

```{r}
#read Birmingham data attributes
BirminghamData <- read_csv("D:\\GIS\\datas\\Birmingham_CDRC_2011_CensusGeodatapack\\tables\\QS102EW_msoa11.csv", na = "n/a")
#rename the columns in BirminghamData
names(BirminghamData)[2:4] <- c("All_usual_residents","Area(Hectares)", "Density_number_of_persons_per_hectare")
```

```{r}
#join the data from census to the in the Birmingham data frame
BirminghamDataMapMSOA <- append_data(BirminghamMapMSOA,BirminghamData, key.shp = "MSOA11CD", key.data = "GeographyCode", ignore.duplicates = TRUE)
```

```{r}
palette1<-scale_fill_continuous(low="white", high="red", "Density_number_of_persons_per_hectare")
labels<-labs(title="Figure 2: Population Density in Birmingham, 2011",x="Longitude", y="Latitude")
ggplot()+geom_sf(mapping = aes(geometry=geometry, fill=Density_number_of_persons_per_hectare),data = BirminghamDataMapMSOA)+theme_minimal()+palette1+labels
```

ArcGIS (figure 1) and R (figure 2) produced two maps, illustrating 2011 census population density in Birmingham. According to Scott and Janikas (2010) and Maindonald and Braun (2010), both ArcGIS and R is an essential software which can not only achieve statistics and analysis of data information but geographic information visualisation. 

ArcGIS, an intuitive graphical user interface (GUI) software, is user-friendly. It has a similar interface with other GUI software so that it is suitable for beginners to explore their data. To create a map, users can click on diverse icons to achieve their intended analysis functions such as clip, buffer and changing the geographic coordinate system. To begin with analysis, Birmingham’s data can be added through clicking and select working directory. Next, those shapefiles and tables in this working directory can be found in ArcCatalog. When files are dragged into the table of contents, Birmingham boundary can be selected by attribute and export as a layer. Then, right-click the boundary layer can join 2011 census data into boundary attribute correctly if the code column is used as a reference. Following this, census data can be geographic analysis via the symbology in layout properties. Here, the graduated colours were break in equal interval. After that, in layout view, key factors such as legend, scale bar mentioned by Krygier and Wood (2016) can be inserted into the final map. 

R also illustrate a similar workflow with ArcGIS, including installing work directory, importing data, filtering and tidying data, analysing data and editing layout, while there are several differences between them. As a command-line interface (CLI) software, R users have to type codes to invoke various functions. It lays a foundation of threshold from users programmed experience otherwise dealing with random errors will be tricky. Before analysis data, vital extra packages should be loaded otherwise some of the code cannot be responded. After that, define names for files and read them in R, following by joining data frame to spatial analysis. To create a map, codes of each layer are input separately then merge.  

Therefore, ArcGIS relies on mouse, while R relies on keyboard. Comparing with R, ArcGIS has a higher degree of freedom in editing layout and create aesthetic maps. As shown in those codes, colour palettes and display styles defined by certain codes, which may restrict beginners’ freedom to design their maps. However, Valero-Mora and Ledesma (2012) highlighted CLI is more accurate statistically than GUI software. Meanwhile, lower speed of data processing is a significant limitation of ArcGIS, contributing frequently crashes and time consumption and R will be faster in comparison. Additionally, ArcGIS has restrictions on operating systems limited MacOS users’ accessed (ESRI, 2017), while R can be installed to various devices freely.

Additionally, after comparing those two maps, the modifiable areal unit problem (MAUP) is a significant issue for 2011 census data. Data are allocated to units of different scale and shapes, for example, LSOAs and MSOAs, contributing different analysis results (Lloyd, 2014). In figure 1, data are located in LSOAs, which including more household per area than MSOAs (ONS, n.d.). Figure 1 has a number of darkest areas than figure 2. Conclusion for figure 1 may be dense population appear in surrounding city centre areas, while conclusion for figure 2 is the population are cluster in a MSOA. This problem is a wider and unsolved yet so that alternative measures to reduce the impact of it can be considered. 

